
\section{Clean Data before Clustering}

The {\it k-mean} algorithm seems relatively simple. Is it actually
useful?  Yes. The beginning of this chapter describes several
scenarios where clustering would be helpful.  Here are some other
applications.  Imagine that you own several pizza stores and want to
open new stores.  You want to decide the new stores' locations.  You
have the budget to open at most five more stores (in this case $k$ is
between 1 to 5).  You want to use the addresses of deliveries to
determine the stores' locations as the centroids of the clusters based
on customers' addresses.  This can be applied to many other scenarios:
As another example, a bank wants to decide where to install four ATM
machines ($k$ is 4) based on customers' home and office addresses.
Both of them can take advantage of the clustering method.

\index{ATM: automatic teller machine}

It is common that the data needs to be ``cleaned'' before sending to a
clustering program.  For example, if you want to decide the locations
of pizza stores, you may want to exclude the customers that are too
far away from the other customers. Maybe you want to exclude the
locations that are, say, more than 10 km away from all the other
customers.  Maybe you also want to consider how often customers order
deliveries and give frequent customers larger weights so that the
stores are closer to these customers.  The same thinking can be
applied to selecting ATM locations: maybe you want to give higher
weights to the customers that have high account balances or use ATM
more often.

Such ``data cleaning'' usually requires human judgement. In the
example of deciding pizza stores' locations, should the customers be
excluded if they are more than 10km away from all the other customers?
Why should it be 10km? Why not 5, or 8, or 15km?  If you want to
include the customers that order deliveries often even though they
live more than 10km away, how do you define ``often''?  Is more than
once per week considered often? Or more than twice per week?  Is one
week too short? Instead, you prefer to consider ordering more than
five times per month?  Doing meaningful data cleaning often requires
knowledge about the business (called {\it domain knowledge} by some
people). You need to know how much it costs to deliver pizza for
customers living at different distances from the store. You also need
to know the traffic conditions: delivering pizza to a customer living
in the center of a city may take longer even though the distance is
shorter.  The domain knowledge is essential understanding what data is
needed, what data should be excluded, and how to interpret the results
from the clustering method.  Using the {\it k-mean} method, or any
machine learning method, is not as simple as throwing data into a
program and get meaningful results effortlessly.

\index{domain knowledge}


% Chapter~\ref{ch:clusterlimit} will explain some other factors for consideration.
